D-Hammer: Practical Proof Automation of Dirac Notation Equations

### 2025-01-29 main.tex
rearranged
.
2025/1/28

% add new commands for comments here
\newcommand{\yx}[1]{\textit{\color{blue}[YX] : #1}}
\newcommand{\lz}[1]{\textit{\color{red}[LZ] : #1}}

@@

@@

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Dirac notation is a fundamental language in quantum computation and quantum information. Recently, the term rewriting system DiracDec~\cite{diracdec} was introduced to automate equational reasoning with Dirac notation, a critical yet time-intensive task. Building upon DiracDec, this work aims to develop a solver optimized for practical applications. Enhancements include an improved typing system, a simplified language and rewriting system, more efficient algorithms, and added support for labeled Dirac notation. Our solver, D-Hammer, is implemented in \CC\ with a Mathematica backend, demonstrating significant improvements in decision-making power and time efficiency.

@@

\input{02techniques.tex}

\input{03tool_eval.tex}

\input{04related_conclusion.tex}

>>>>

% add new commands for comments here
\newcommand{\yx}[1]{\textit{\color{blue}[YX] : #1}}
\newcommand{\gb}[1]{\textit{\color{green}[GB] : #1}}
\newcommand{\lz}[1]{\textit{\color{red}[LZ] : #1}}

@@

\usepackage{algorithm}
\usepackage{algpseudocode}

@@

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Labelled Dirac notation is a formalism commonly used by physicists to
represent many-body quantum systems and computer scientists to assert
properties of quantum programs. Labelled Dirac notation is supported
by a rich equational theory for proving equality between expressions
of the language. These proofs are typically carried on pen-and-paper,
and can be exceedingly long and error-prone. We introduce D-Hammer,
the first tool to supported automated equational proof for labelled
Dirac notation. The salient features of D-Hammer are: an expressive,
higher-order, dependently-typed language for labelled Dirac notation;
an efficient algorithm to normalize expressions (modulo associativity
and commutativity); and an optimized \CC\ implementation. We evaluate
the implementation on representative examples from plain and labelled
Dirac notation. In the case of plain Dirac notation, we show that our
implementation significantly outperforms DiracDec (Xu et al, POPL'25).

@@

\input{02language_design.tex}

\input{03labelled.tex}

\input{04efficient_alg.tex}

\input{05tool_eval.tex}

\input{06related_conclusion.tex}
